Numerical investigation of viscous flow instabilities through heterogeneous porous media. Advantages and limitations of unstructured and adaptive meshes in 2D and 3D domains under different mobility scenarios.

### 2017-12-07 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
change group name

%%%%%%%%%%%%%
\subsection{Capturing Flow Dynamics and Fingering Growth: Impact of Mesh Resolution}\label{section:results_hete_fix_adapt}    
%\subsection{Comparison of fixed and adaptive mesh under heterogeneous domain schemes and the same mobility ratio.}
In most numerical simulations involving geofluid dynamics, local geometric constraints (\eg faults, fractures etc) and spatial multi-scale variability of flow properties are often ignored as the underlying computational mesh grid is too coarse to reliably represent any of these features. Whilst structured grids often struggle to conform to complex domain boundaries with consistent mesh connectivity, unstructured mesh techniques often relax cells' neighbourhood relationship constraints by dividing the domain into polytopes in which elements share adjacent faces. This leads to mesh grids that conform to the domain topography and can make the best use of state-of-the-art self-adaptive computational methods.

@@

\medskip
Traditional numerical methods often rely on fixed mesh with sufficient resolution throughout the domain to capture specific flow dynamics (\eg fluid instabilities, flow recirculation, heat and mass transfers etc), however computational resources may make it prohibitive for complex geometries and highly heterogeneous properties.  Mesh adaptivity methods have been extensively used by the computational geofluid dynamic community to help capturing detailed flow dynamics, compositional non-equilibrium fluid displacement and solid-fluid interactions \citep{pluszny_2007,pietro_2014,su_2016,melnikova_2016}. In these methods, the mesh grid is continuously modified (\ie adjustments of the number and distribution of the degrees of freedom to reduce solution error) to focus resolution where is necessary as the simulation evolves in time and space. There are four main families of mesh adaptivity methods: adaptive mesh refinement (AMR), edge and face element manipulations whereas keeping the order of the element basis function fixed (h-adaptivity), mesh deformation (r-adaptivity) and changes of element basis function order (p-adaptivity). Detailed description and analysis of mesh adaptivity methods are beyond the scope of this manuscript but can be readily found in \citet{lo_book} \citep[see also][]{plewa_book,frey_book}.

>>>>

%%%%%%%%%%%%%
\subsection{Capturing Flow Dynamics and Fingering Growth: Impact of Mesh Resolution}\label{section:results_hete_fix_adapt}    
%\subsection{Comparison of fixed and adaptive mesh under heterogeneous domain schemes and the same mobility ratio.}
In most numerical simulations involving geo-fluid dynamics, local geometric constraints (\eg faults, fractures etc) and spatial multi-scale variability of flow properties are often ignored as the underlying computational mesh grid is too coarse to reliably represent any of these features. Whilst structured grids often struggle to conform to complex domain boundaries with consistent mesh connectivity, unstructured mesh techniques often relax cells' neighbourhood relationship constraints by dividing the domain into polytopes in which elements share adjacent faces. This leads to mesh grids that conform to the domain topography and can make the best use of state-of-the-art self-adaptive computational methods.

@@

\medskip
Traditional computational geo-fluid dynamics (CGFD) models often rely on fixed mesh with sufficient resolution throughout the domain to capture specific flow dynamics (\eg fluid instabilities, flow recirculation, heat and mass transfers etc), however computational resources may make it prohibitive for complex geometries and highly heterogeneous properties.  Mesh adaptivity methods have been extensively used by the computational geo-fluid dynamic community to help capturing detailed flow dynamics, compositional non-equilibrium fluid displacement and solid-fluid interactions \citep{pluszny_2007,pietro_2014,su_2016,melnikova_2016}. In these methods, the mesh grid is continuously modified (\ie adjustments of the number and distribution of the degrees of freedom to reduce solution error) to focus resolution where is necessary as the simulation evolves in time and space. There are four main families of mesh adaptivity methods: adaptive mesh refinement (AMR), edge and face element manipulations whereas keeping the order of the element basis function fixed (h-adaptivity), mesh deformation (r-adaptivity) and changes of element basis function order (p-adaptivity). Detailed description and analysis of mesh adaptivity methods are beyond the scope of this manuscript but can be readily found in \citet{lo_book} \citep[see also][]{plewa_book,frey_book}.
